---
title: "Calibration result on sibling data: random genotype"
output: html_document
date: '2024-01-31'
---

Description: Simulation result for calibration. The simulation model is the same as https://yunqiyang0215.github.io/simulation-calibrated-mr/ukb_calibration.html. Instead of using real genotype from UKB, we simulate genotype using UKB allele frequency in each simulation replicates. Therefore, genotype is random in this case. 


```{r}
res = readRDS("/project2/mstephens/yunqiyang/calibrated_mr/simulation_sibling/linear202403/res.rds")
res.sub = readRDS("/project2/mstephens/yunqiyang/calibrated_mr/simulation_sibling/linear202403/res.sub.rds")
```

```{r}
combine_single_snp_stat = function(cond_indx, snp_indx, data.list, column_names){
  res = lapply(cond_indx, function(ind) data.list[[ind]][snp_indx, column_names])
  dat_combined = do.call(rbind, res)
  return(dat_combined)
}
```

### 1. Check correlations:

#### 1.1 Phenotype correlation when r1 = 0 and r2 = 0.1

```{r fig.width=6, fig.height=10}
par(mfrow = c(4,2))
p = c(10, 100)
sig = c(1, 10)
for (i in 1:length(p)){
  for (j in 1:length(sig)){
    indx = which(res.sub$sim_data.sig == sig[j] & res.sub$sim_effect.p == p[i])
    corr1 = sapply(indx, function(i) cor(res.sub$sim_data.dat[[i]]$pheno1[, 1], res.sub$sim_data.dat[[i]]$pheno2[, 1]))
    corr2 = sapply(indx, function(i) cor(res.sub$sim_data.dat[[i]]$pheno1[, 2], res.sub$sim_data.dat[[i]]$pheno2[, 2]))
    hist(corr1, main = paste0("Correlation pheno1:", "p=", p[i], ",sigma=", sig[j]), xlab = "")
    hist(corr2, main = paste0("Correlation pheno2:", "p=", p[i], ",sigma=", sig[j]), xlab = "")
  }
}




```



### 2. Compare calibrated estimator vs. raw estimator (Exposure trait)

#### 2.1 Parameters: p = 10 and sig = 1

```{r fig.width=8, fig.height=20}

par(mfrow = c(5,2))
r1 = unique(res$sim_data.r1)

for (i in 1:5){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 10 & res$sim_data.sig == 1), i, data.list = res$fit.sumstat, column_names = c("Y1.cali", "Y1.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 10 & res$sim_data.sig == 1), i, data.list = res$fit.sumstat, column_names = c("Y1.cali.var", "Y1.raw.var")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,

          ylim = c(0, max(dat_estimator_var, empirical_var)+0.005),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}
```


#### 2.2 Parameters: p = 10 and sig = 10

```{r fig.width=8, fig.height=20}

par(mfrow = c(5,2))
r1 = unique(res$sim_data.r1)
for (i in 1:5){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 10 & res$sim_data.sig == 10), i, data.list = res$fit.sumstat, column_names = c("Y1.cali", "Y1.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 10 & res$sim_data.sig == 10), i, data.list = res$fit.sumstat, column_names = c("Y1.cali.var", "Y1.raw.var")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,

          ylim = c(0, max(dat_estimator_var, empirical_var)+0.005),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}
```

 
#### 2.3 Parameters: p = 100 and sig = 1

```{r fig.width=8, fig.height=20}

par(mfrow = c(5,2))
r1 = unique(res$sim_data.r1)
for (i in 1:5){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 100 & res$sim_data.sig == 1), i, data.list = res$fit.sumstat, column_names = c("Y1.cali", "Y1.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$sim_effect.p == 100 & res$sim_data.sig == 1), i, data.list = res$fit.sumstat, column_names = c("Y1.cali.var", "Y1.raw.var")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,

          ylim = c(0, max(dat_estimator_var, empirical_var)+0.005),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}
```



