---
title: "Check variance reduction empirically"
author: "Yunqi Yang"
date: "4/4/2023"
output: html_document
---

#### Description:

check variance reduction using estimated mean and variance of score functions. And compute variance reduction empirically under Gaussian x1, x2 and binary x1, x2. Binary x1, x2 are obtained by first simulating from joint Gaussian and then transformed to binary data based on threshold $x>0$. 

**Theoretical value of variance reduction**
$$
VR = \frac{(1-r)(1-\rho)}{2}\frac{\sigma^2}{\sigma^2+(1-r^2)\beta_2^2}
$$

**Conclusion**: the patterns for variance reduction between Gaussian x1, x2 and binary x1, x2 are very similar. 

```{r  }
compute_var_reduction <- function(sigma2, rho, r, b2){
  vr <- (1-r)*(1-rho)/2 * sigma2 / (sigma2 + (1-r^2)*b2^2)
  return(vr)
}
```


### 1. Vary correlation 

data replicates: 5000

n = 5e3; N = 5e4; b = c(1,3,0.1); sigma = 1

```{r}
gaussian <- readRDS("./data/simulation202304/gaussian_r.rds")
binary <- readRDS("./data/simulation202304/binary_r.rds")
```

```{r}
rs <- c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8)
est.r = rep(NA, length(binary))

var1 <- matrix(NA, ncol = 2, nrow = length(gaussian))
var2 <- matrix(NA, ncol = 2, nrow = length(binary))
colnames(var1) <- c("gamma.tilde", "gamma.raw")
colnames(var2) <- c("gamma.tilde", "gamma.raw")


for (i in 1:length(gaussian)){
  var1[i, ] <- apply(gaussian[[i]][,c(1,2)], MARGIN = 2, function(x) var(x))
}

for (i in 1:length(binary)){
  var2[i, ] <- apply(binary[[i]][, c(1,2)], MARGIN = 2, function(x) var(x))
  est.r[i] = mean(binary[[i]][,4])
}


```


```{r}
### Mean variance of calibrated estimator
mean_var_gaussian <- rep(NA, length(gaussian))
mean_var_binary <- rep(NA, length(binary))
for (i in 1:length(gaussian)){
  mean_var_gaussian[i] = mean(gaussian[[i]][,3])
}

for (i in 1:length(binary)){
  mean_var_binary[i] = mean(binary[[i]][,3])
}
```




```{r}
plot(rs, var1[,1], col = 2, ylim = c(0, 0.0050), xlab = "correlation between x1 and x2", ylab = "estimated variance", pch = 19)
points(rs, var1[,2], pch = 19)


points(est.r, var2[,1], col = 2, pch = 17, xlab = "correlation between x1 and x2", ylab = "estimated variance")
points(est.r, var2[,2], pch = 17)
legend("topleft", legend = c("Gaussian: tau.tilde", "Gaussian: tau.raw", 
                             "Binary: tau.tilde","Binary: tau.raw", "Gaussian: mean(var.hat) of cali-estimator", "Binary: mean(var.hat) of cali-estimator"), col = c(2,1, 2,1, 1, 4), pch = c(19, 19, 17, 17, 1, 1))

points(rs, mean_var_gaussian)
points(est.r, mean_var_binary, col = 4)
```

```{r}
grid = seq(-0.8, 0.8, length.out = 100)
theo = rep(NA, length(grid))
for (i in 1:length(grid)){
  theo[i] <- compute_var_reduction(sigma2 = 1, rho = 0.1, grid[i], b2 = 0.1)
}
```

```{r}
plot(rs, 1-var1[,1]/var1[,2], xlab = "correlation between x1 and x2", ylab = 'proportion of variance reduction', pch = 19, col = "red")

points(est.r, 1-var2[,1]/var2[,2], pch = 17, col = "blue")
lines(grid, theo, lty = 2)
legend("topright", legend = c("Gaussian", "Binary"), col = c("red","blue"), pch = c(19, 17))
```



### 2. Vary rhos: 1000 data replicates

b <- c(1,3,0.1); sigma <- 1; r <- 0; N <- 1e4

```{r}
gaussian <- readRDS("./data/simulation202304/gaussian_rho.rds")
binary <- readRDS("./data/simulation202304/binary_rho.rds")
```

```{r}
rhos <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
var1 <- matrix(NA, ncol = 2, nrow = length(rhos))
var2 <- matrix(NA, ncol = 2, nrow = length(rhos))
colnames(var1) <- c("gamma.tilde", "gamma.raw")
colnames(var2) <- c("gamma.tilde", "gamma.raw")

for (i in 1:length(rhos)){
  var1[i, ] <- apply(gaussian[[i]][,c(1,2)], MARGIN = 2, function(x) var(x))
  var2[i, ] <- apply(binary[[i]][,c(1,2)], MARGIN = 2, function(x) var(x))
}

```

```{r}
### Mean variance of calibrated estimator
mean_var_gaussian <- rep(NA, 9)
mean_var_binary <- rep(NA, 9)
for (i in 1:9){
  mean_var_gaussian[i] = mean(gaussian[[i]][,3])
}

for (i in 1:9){
  mean_var_binary[i] = mean(binary[[i]][,3])
}
```

```{r}
plot(rhos, var1[,1], col = 2, ylim = c(0, 0.0090), xlab = "rho: n/N", ylab = "estimated variance", pch = 19)
points(rhos, var1[,2], pch = 19)

points(rhos, var2[,1], col = 2, pch = 17)
points(rhos, var2[,2], pch = 17)
points(rhos, mean_var_gaussian)
points(rhos, mean_var_binary, col = 4)

legend("topright", legend = c("Gaussian: tau.tilde", "Gaussian: tau.raw", 
                             "Binary: tau.tilde","Binary: tau.raw", "Gaussian: mean(var.hat) of cali-estimator", "Binary: mean(var.hat) of cali-estimator"), col = c(2,1, 2,1, 1, 4), pch = c(19, 19, 17, 17, 1, 1))


```


```{r}
grid = seq(0, 1, length.out = 100)
theo = rep(NA, length(grid))
for (i in 1:length(grid)){
  theo[i] <- compute_var_reduction(sigma2 = 1, rho = grid[i], r = 0, b2 = 0.1)
}

```

```{r}
plot(rhos, 1-var1[,1]/var1[,2], xlab = "rho: n/N", ylab = 'proportion of variance reduction', pch = 19, col = "red",
      xlim = c(0, 1), ylim = c(0, 0.6))
points(rhos, 1-var2[,1]/var2[,2], pch = 17, col = "blue")
#points(rhos, 1-mean_var_gaussian/var1[,2], col = 1)
#points(rhos, 1-mean_var_binary/var2[,2], col = 4)
lines(grid, theo, lty = 2)
legend("topright", legend = c("Gaussian", "Binary"), col = c("red","blue"), pch = c(19, 17))
```





