---
title: "Calibration result: random genotype"
output: html_document
date: '2024-01-31'
---

Description: Simulation result for calibration. The simulation model is the same as https://yunqiyang0215.github.io/simulation-calibrated-mr/ukb_calibration.html. Instead of using real genotype from UKB, we simulate genotype using UKB allele frequency in each simulation replicates. Therefore, genotype is random in this case. 


```{r}
res = readRDS("/project2/mstephens/yunqiyang/calibrated_mr/mr_dsc20230727/res.rds")
```

```{r}
combine_single_snp_stat = function(cond_indx, snp_indx, data.list, column_names){
  res = lapply(cond_indx, function(ind) data.list[[ind]][snp_indx, column_names])
  dat_combined = do.call(rbind, res)
  return(dat_combined)
}
```

### 1. Compare calibrated estimator vs. raw estimator (Exposure trait)

```{r fig.width=8, fig.height=40}

par(mfrow = c(10,2))
r1 = unique(res$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y1.cali", "Y1.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y1.cali.var", "Y1.var.raw")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          #ylim = c(0, 0.06),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}
```


### 2. Compare variance reduction (Exposure trait)

```{r fig.width=14, fig.height=6}

par(mfrow = c(2,5))
r1 = unique(res$sim_data.r1)
p = 10

for (i in 1:p){
  
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y1.cali", "Y1.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r1), function(k) {
  stats <- combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y1.cali.var", "Y1.var.raw"))
  variance_reduction <- (stats[,"Y1.var.raw"] - stats[,"Y1.cali.var"]) / stats[,"Y1.var.raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r1",
          ylab = "Variance Reduction",
          ylim = c(min(dat_variance_reduction) - 0.01, max(dat_variance_reduction) + 0.1),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}
```



 
 
 
### 3. Compare calibrated estimator vs. raw estimator (Outcome trait)

```{r fig.width=8, fig.height=40}

par(mfrow = c(10,2))
r1 = unique(res$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali", "Y2.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali.var", "Y2.var.raw")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          #ylim = c(0, 0.06),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}
```


```{r fig.width=8, fig.height=40}

par(mfrow = c(10,2))
r2 = unique(res$sim_data.r2)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r2), function(k) combine_single_snp_stat(which(res$sim_data.r2 == r2[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali", "Y2.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r2",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r2)), labels=r2)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r2), function(k) combine_single_snp_stat(which(res$sim_data.r2 == r2[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali.var", "Y2.var.raw")))
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          #ylim = c(0, 0.06),
          xlab = "r2",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- r2
  axis(1, at=seq(1.5, by=2, length.out=length(r2)), labels=r2)
}
```



### 4. Compare variance reduction (Outcome trait)

```{r fig.width=14, fig.height=6}

par(mfrow = c(2,5))
r1 = unique(res$sim_data.r1)
p = 10

for (i in 1:p){
  
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali", "Y2.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r1), function(k) {
  stats <- combine_single_snp_stat(which(res$sim_data.r1 == r1[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali.var", "Y2.var.raw"))
  variance_reduction <- (stats[,"Y2.var.raw"] - stats[,"Y2.cali.var"]) / stats[,"Y2.var.raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r1",
          ylab = "Variance Reduction",
          ylim = c(min(dat_variance_reduction) - 0.01, max(dat_variance_reduction) + 0.1),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}
```

```{r fig.width=14, fig.height=6}

par(mfrow = c(2,5))
r2 = unique(res$sim_data.r2)
p = 10

for (i in 1:p){
  
  snp_dat_list = lapply(1:length(r2), function(k) combine_single_snp_stat(which(res$sim_data.r2 == r2[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali", "Y2.raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r2), function(k) {
  stats <- combine_single_snp_stat(which(res$sim_data.r2 == r2[k] & res$fit.method == "calibrated"), i, data.list = res$fit.sumstat, column_names = c("Y2.cali.var", "Y2.var.raw"))
  variance_reduction <- (stats[,"Y2.var.raw"] - stats[,"Y2.cali.var"]) / stats[,"Y2.var.raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r2",
          ylab = "Variance Reduction",
          ylim = c(min(dat_variance_reduction) - 0.01, max(dat_variance_reduction) + 0.1),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r2)), labels=r2)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}
```

### 5. Compare results with G & Gpa model
```{r}
indx1 = which(res$sim_data.r1 == 0.5 & res$fit.method == "calibrated")
indx2 = which(res$sim_data.r1 == 0.5 & res$fit.method == "calibrated_model_G")
```

```{r}
print(res$fit.sumstat[[indx1[1]]])
print(res$fit.sumstat[[indx2[1]]])
```


```{r}
indx1 = which(res$sim_data.r1 == -0.2 & res$fit.method == "calibrated")
indx2 = which(res$sim_data.r1 == -0.2 & res$fit.method == "calibrated_model_G")
```

```{r}
print(res$fit.sumstat[[indx1[1]]])
print(res$fit.sumstat[[indx2[1]]])
```
