---
title: "Variance reduction: binomial x1, x2"
author: "Yunqi Yang"
date: "3/28/2023"
output: html_document
---

## Description:

simulation to check variance reduction under binomial x1 and x2. x1 and x2 are first simulated from gaussian and then transformed to binomial distribution. 

The calibrated estimator is calculated using:
$$
\tilde\tau= \hat\beta_1-\hat\beta_2-v_{12}v_{22}^{-1}(\hat\alpha_1'-\hat\alpha_1),
$$
where $v_{12}$ and $v_{22}$ are theoretical values. $\hat\alpha_1$ is the estimate from larger external data under mis-specified model. 


### 1. Theoretical value of variance reduction (under x1, x2 joint gaussian case)

$$
VR = \frac{(1-r)(1-\rho)}{2}\frac{\sigma^2}{\sigma^2+(1-r^2)\beta_2^2}
$$
```{r}
compute_var_reduction <- function(sigma2, rho, r, b2){
  vr <- (1-r)*(1-rho)/2 * sigma2 / (sigma2 + (1-r^2)*b2^2)
  return(vr)
}
```

### 2. Theoretical vs. simulation: Change correlation 

Script available: 

n = 5e3; N = 5e4; b = c(1,3,0.1); sigma = 1

$\tilde\tau$ is calculated using the equation under Gaussian case, and we use estimated empirical correlation for $r$, the correlation parameter. 
```{r}
res <- readRDS("./data/sim_binomial/binomial_r.rds")
```

```{r}
rs <- c(-0.8, -0.6, -0.4, -0.2, 0, 0.2, 0.4, 0.6, 0.8)
est_var <- matrix(NA, ncol = 3, nrow = length(res))
colnames(est_var) = c("gamma.tilde", "gamma.raw", "est.r")
for (i in 1:length(res)){
  est_var[i, 1:2] <- apply(res[[i]][,1:2], MARGIN = 2, function(x) var(x))
  est_var[i, 3] <- mean(res[[i]][, 3])
}
```

```{r}
theo_vr <- rep(NA, length(res))
for (i in 1:length(res)){
  theo_vr[i] <- compute_var_reduction(sigma2 = 1, rho = 0.1, r = mean(res[[i]][, 3]), b2 = 0.1)
}

```

```{r}
plot(est_var[,3], 1-est_var[,1]/est_var[,2], xlab ="estimated correlation between x1 and x2", ylab = "Variance reduction proportion")
points(est_var[,3], theo_vr, col = "red")
legend("topright", legend = c("empirical", "theoretical"), pch = 1, col = c(1,2))
```
**Comments**: we should use empirically estimated $\tilde\tau$ (by estimating the empirical covariance of score functions...), instead of using Gaussian equation to get $\tilde\tau$. 


### 3. Theoretical vs. simulation: Change rho

Here I fixed N = 5e4, and vary n based on different $\rho$ values. 

Simulation script: 

n = 5e3; b = c(1,3,0.1); sigma = 1; r = 0

```{r}
res <- readRDS("./data/sim_binomial/binomial_rho.rds")
```

```{r}
rhos <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1)
est_var <- matrix(NA, ncol = 3, nrow = length(res))
colnames(est_var) = c("gamma.tilde", "gamma.raw", "est.r")
for (i in 1:length(res)){
  est_var[i, 1:2] <- apply(res[[i]][,1:2], MARGIN = 2, function(x) var(x))
  est_var[i, 3] <- mean(res[[i]][, 3])
}
```

```{r}
theo_vr <- rep(NA, length(res))
for (i in 1:length(res)){
  theo_vr[i] <- compute_var_reduction(sigma2 = 1, rho = rhos[i], r = mean(res[[i]][, 3]), b2 = 0.1)
}
```

```{r}
plot(rhos, 1-est_var[,1]/est_var[,2], xlab ="rho (n/N)", ylab = "Variance reduction proportion", ylim = c(0, 0.5))
points(rhos, theo_vr, col = "red")
legend("topright", legend = c("empirical", "theoretical"), col = c(1,2), pch = 1)
```



