---
title: "Calibration result on UKB genotype"
author: "Yunqi Yang"
date: "02/01/2024"
output: html_document
---

### Simulation model and setting:

1. $p = 10$

2. $b, b_p$ are fixed. 

3. Model to simulate parents' phenotype:

$$
\begin{split}
b^p&\sim N(0, 1),\quad \epsilon\sim N(0, 1)\\
Y^p &= G^pb^p+\epsilon
\end{split}
$$

4. Model to simulate children's exposure & outcome traits: $\beta$ denotes the causal effect. $T$ denotes transmitted allele matrix, and $NT$ denotes non-transmitted allele matrix. 
$$
\begin{split}
\beta & =1\\
b&\sim N(0, 1), \quad \epsilon\sim N(0, 1)\\
Y_1 &= Tb + r_1(Y^m+Y^f) + \epsilon\\
Y_2 &= \beta Y_1 + r_2(Y^m + Y^f) + \epsilon
\end{split}
$$
If we write $G^p$ as $T+NT$, then we have:

$$
\begin{split}
Y1 = \sum_j^p\Big[(b^p_jr_1+b_j)T_j+r_1b_j^pNT_j\Big]+\epsilon^p+\epsilon
\end{split}
$$
If we assume $y=\beta_0 + \beta_1x_1 + \beta_2x_2 +\epsilon$ is the true model. $[x_1, x_2]$ follows normal distribution and $\epsilon\sim N(0, \sigma^2)$. The theoretical variance reduction under Gaussian X is as below, where $\rho$ is the ratio of sample size between internal complete data and external incomplete data, and $r$ is the correlation between $x_1$ and $x_2$. 

$$
\begin{split}
VR = \frac{(1-r)(1-\rho)}{2}\frac{\sigma^2}{\sigma^2+(1-r^2)\beta_2^2}
\end{split}
$$


```{r}
#res1 = readRDS("/project2/mstephens/yunqiyang/calibrated_mr/simulation-calibrated-mr/data/simulation202401/res.rds") # previous simulation result
res = readRDS("/project2/mstephens/yunqiyang/calibrated_mr/sim202401/res.rds") # current result
res.Y1 = res[[1]]
```

```{r}
combine_single_snp_stat = function(cond_indx, snp_indx, data.list, column_names){
  res = lapply(cond_indx, function(ind) data.list[[ind]][snp_indx, column_names])
  dat_combined = do.call(rbind, res)
  return(dat_combined)
}
```



### 1. Multiple effects without resampling: 

```{r eval = FALSE}
set.seed(1)
p = 10
b = matrix(rnorm(p, sd = 1), ncol = 1, nrow = p)
bp = matrix(rnorm(p, sd = 1), ncol = 1, nrow = p)
```

```{r fig.width=8, fig.height=40}
par(mfrow = c(10,2))
r1 = unique(res.Y1$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw")))
  
  
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          ylim = c(0, max(dat_estimator_var)+0.01),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}

```

```{r fig.width=14, fig.height=6}
par(mfrow = c(2,5))
r1 = unique(res.Y1$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))
  

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r1), function(k) {
  stats <- combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == FALSE), i, data.list = res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw"))
  variance_reduction <- (stats[, "var_tau_raw"] - stats[,"var_tau_cal"]) / stats[, "var_tau_raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r1",
          ylab = "Variance Reduction",
          ylim = c(0, max(dat_variance_reduction) + 0.1),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}

```


### 2. Multiple effects with resampling 

```{r eval = FALSE}
set.seed(1)
p = 10
b = matrix(rnorm(p, sd = 1), ncol = 1, nrow = p)
bp = matrix(rnorm(p, sd = 1), ncol = 1, nrow = p)
```

```{r fig.width=8, fig.height=40}
par(mfrow = c(10,2))
r1 = unique(res.Y1$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == TRUE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == TRUE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw")))
  
  
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          ylim = c(0, max(dat_estimator_var)+0.01),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}

```

```{r fig.width=14, fig.height=6}
par(mfrow = c(2,5))
r1 = unique(res.Y1$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == TRUE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))
  

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r1), function(k) {
  stats <- combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "multiple" & res.Y1$sim_data.resample == TRUE), i, data.list = res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw"))
  variance_reduction <- (stats[, "var_tau_raw"] - stats[,"var_tau_cal"]) / stats[, "var_tau_raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r1",
          ylab = "Variance Reduction",
          ylim = c(-0.05, max(dat_variance_reduction) + 0.1),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}

```


### 3. Single effect:

```{r eval = FALSE}

set.seed(1)
b = bp = matrix(rep(0, p), ncol = 1, nrow = p)
b[1,1] = rnorm(1)
bp[1,1] = rnorm(1)

```

```{r fig.width=8, fig.height=40}
par(mfrow = c(10,2))
r1 = unique(res.Y1$sim_data.r1)

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "single" &
                                                                                  res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))

  dat_estimator <- do.call(cbind, snp_dat_list)
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator))
  boxplot(dat_estimator,
          #main = paste0("Exposure trait, r1=", r1[k]),
          xlab = "r1",
          ylab = "(un)calibrated estimator",
          col = colors,
          outpch = 20, # This changes the outlier point type to solid circle
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator), colMeans(dat_estimator), pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical mean"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)

  
  ### Plot variance
  empirical_var = apply(dat_estimator, 2, var)
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "single"
                                                                                & res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw")))
  
  
  dat_estimator_var <- do.call(cbind, snp_dat_list)
    
  colors <- rep(c("lightblue", "gold1"), length.out = ncol(dat_estimator_var))
  boxplot(dat_estimator_var,
          #main = paste0("Exposure trait, r1=", r1[k]),
          ylim = c(0, max(dat_estimator_var)+0.01),
          xlab = "r1",
          ylab = "Variance of (un)calibrated estimator",
          col = colors,
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  points(1:ncol(dat_estimator_var), empirical_var, pch = 4, cex = 0.8, col = "red2")
  legend("topright",
       legend = c("Calibrated Estimator", "Uncalibrated Estimator", "Empirical variance"),
       fill = c("lightblue", "gold1", NA),  
       pch = c(NA, NA, 4),  
       col = c(NA, NA, "red2"),  
       cex = 0.6)  
  r_values <- c(-0.20, -0.10, 0.00, 0.05, 0.10, 0.50) 
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
}

```




```{r fig.width=14, fig.height=6}
par(mfrow = c(2,5))
r1 = unique(res.Y1$sim_data.r1)
p = 10

for (i in 1:p){
  snp_dat_list = lapply(1:length(r1), function(k) combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "single" & res.Y1$sim_data.resample == FALSE), i, data.list =res.Y1$calibration.Y1_cali, column_names = c("tau_cal", "tau_raw")))
  

  dat_estimator <- do.call(cbind, snp_dat_list)
  empirical_var = apply(dat_estimator, 2, var)
  
  # Separate var_cali and var_raw values
  var_cali <- empirical_var[c(TRUE, FALSE)]
  var_raw <- empirical_var[c(FALSE, TRUE)]
  empirical_var_reduction <- (var_raw - var_cali) / var_raw

  
  # Calculate variance reduction
  snp_dat_list <- lapply(1:length(r1), function(k) {
  stats <- combine_single_snp_stat(which(res.Y1$sim_data.r1 == r1[k] & res.Y1$sim_effect.effect_type == "single" & res.Y1$sim_data.resample == FALSE), i, data.list = res.Y1$calibration.Y1_cali, column_names = c("var_tau_cal", "var_tau_raw"))
  variance_reduction <- (stats[, "var_tau_raw"] - stats[,"var_tau_cal"]) / stats[, "var_tau_raw"]
  return(variance_reduction)
})

  # Combine the data
  dat_variance_reduction <- do.call(cbind, snp_dat_list)
  
  # Boxplot for variance reduction
  boxplot(dat_variance_reduction,
          xlab = "r1",
          ylab = "Variance Reduction",
          ylim = c(0.2, max(dat_variance_reduction) + 0.2),
          col = "lightblue",  
          outpch = 20, 
          outcex = 0.5,
          xaxt = "n")
  

  points(1:length(empirical_var_reduction), empirical_var_reduction, pch = 4, cex = 0.8, col = "red2")
  axis(1, at=seq(1.5, by=2, length.out=length(r1)), labels=r1)
  legend("topright", 
       legend = c("Simulated Variance Reduction", "Empirical Variance Reduction"),
       col = c("lightblue", "red2"),
       pch = c(15, 4),
       cex = 0.8)
}

```










